Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-07-29 emsoft-2019/camera_ready/experiments.tex
My pass on the final version.

We evaluate a prototype implementation of the algorithm in~\autoref{lst:alg}
on three examples. For the first two, we apply the complete pipeline (design and memory optimization)
which returns an end-to-end robust controller. With the third example, we evaluate
the scalability of our approach when the number of regions and hyperplanes are in
the order of tens of thousands.

>>>>

We evaluate a prototype implementation of the algorithm in~\autoref{lst:alg}
on three examples.\footnote{
 Our implementation is available online at \url{https://github.com/rospoly/rmpc-daisy}.
}
For the first two, we apply the complete pipeline (design and memory optimization)
which returns an end-to-end robust controller. With the third example, we evaluate
the scalability of our approach when the number of regions and hyperplanes are in
the order of tens of thousands.

### 2019-07-29 emsoft-2019/camera_ready/finite_precision.tex
My pass on the final version.

The controller has to be stored and evaluated on a microcontroller in
finite precision, since infinite-precision real values cannot be stored on
today's digital computers.
The standard choice for low-power platforms is fixed-point arithmetic. Unlike
floating-point arithmetic~\cite{ieee75408}, fixed-point arithmetic can be implemented efficiently
without complex hardware support using only integer operations. It requires,
however, more compilation effort to determine certain operations statically at
compile time (that the floating-point hardware unit performs dynamically).

@@

% introduces roundoff errors, these can be tracked with a dataflow analysis
Finite precision unavoidably introduces roundoff errors, which can accumulate
during the course of a computation.
A sound static roundoff error analysis computes a guaranteed upper bound on the
error of the result by tracking worst-case errors at every operation. Given a
word length for each value, several tools, including the tool Daisy which we use, automatically determine the number of
integer bits needed and compute roundoff errors using a dataflow analysis~\cite{Daisy,Gappa}.
They track real-valued ranges at every intermediate operation. These ranges
determine the fixed-point formats and thus also the individual worst-case roundoff errors,
which the analyses track separately from the ranges. To ensure guaranteed
error bounds, both the ranges and errors are computed using interval~\cite{intervals} or
affine arithmetic~\cite{affineArithmetic} which compute sound enclosures. 
Affine arithmetic tracks linear correlations between variables, so for linear expressions
the computed ranges are as tight as possible (nonlinear arithmetic
leads to over-approximations).
% A standard technique to increase accuracy of the computed errors is interval
% subdivision, which divides the input domain into equally-sized subdomains and
% runs roundoff error analysis separately for each subdomain. The
% over-approximations on each subdomain tend to be smaller, which increases the
% overall tightness of range and error bounds.

@@

In this work, we use the tool Daisy which implements roundoff error
analysis~\cite{Daisy} and mixed-precision tuning~\cite{DaisyTuning} for fixed-point arithmetic, and is
available as open-source. Fixed-point arithmetic implementations can choose different
rounding modes; here we consider truncation, which is more efficient than
rounding.

>>>>

We now discuss finite-precision issues in the implementation of a controller.
% The controller has to be stored and evaluated on a microcontroller in finite precision.
%  since infinite-precision real values cannot be stored on today's digital computers.
The standard choice for low-power platforms is fixed-point arithmetic. Unlike
floating-point arithmetic~\cite{ieee75408}, fixed-point arithmetic can be implemented efficiently
without complex hardware support using only integer operations. It requires,
however, more compilation effort to determine certain operations statically at
compile time that the floating-point hardware unit performs dynamically.

@@

% introduces roundoff errors, these can be tracked with a dataflow analysis
Finite precision introduces roundoff errors, which can accumulate
during the course of a computation.
A sound static roundoff error analysis computes a guaranteed upper bound on the
error of the result by tracking worst-case errors at every operation. Given a
word length for each value, several tools, including the tool Daisy which we use, automatically determine the number of
integer bits needed and compute roundoff errors using a dataflow analysis~\cite{Daisy,Gappa}.
They track real-valued ranges at every intermediate operation. These ranges
determine the fixed-point formats and thus also the individual worst-case roundoff errors,
which the analyses track separately from the ranges. To ensure guaranteed
error bounds, both the ranges and errors are computed using interval~\cite{intervals} or
affine arithmetic~\cite{affineArithmetic}, which compute sound enclosures. 
Affine arithmetic tracks linear correlations between variables, so for linear expressions
the computed ranges are as tight as possible (nonlinear arithmetic
leads to over-approximations).
% A standard technique to increase accuracy of the computed errors is interval
% subdivision, which divides the input domain into equally-sized subdomains and
% runs roundoff error analysis separately for each subdomain. The
% over-approximations on each subdomain tend to be smaller, which increases the
% overall tightness of range and error bounds.

@@

In this work, we use the open-source tool Daisy which implements roundoff error
analysis~\cite{Daisy} and mixed-precision tuning~\cite{DaisyTuning} for fixed-point arithmetic.
% and is available as open-source. 
Fixed-point arithmetic implementations can choose different
rounding modes; here we consider truncation, which is more efficient than rounding.

### 2019-07-29 emsoft-2019/camera_ready/intro.tex
My pass on the final version.

To the best of our knowledge, our algorithm is the first to automatically and soundly
synthesize a suitable fixed-point precision for explicit robust MPC controllers,
and the first to consider mixed-precision in this context.

@@

For smaller examples, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few hours.
In absolute terms, the memory saving corresponds to 21KB over a uniform
precision implementation on our largest benchmark. %\Mahmoud{It should be noticed that these savings are not affecting the run-time. Synthesis of the controller and the corresponding precisions are all performed offline.}
\Mahmoudd{We note that EMPC is a class of controllers that is mainly appropriate for implementation over processors with limited computational power and memory. Typical memory of these processors is in the order of tens of kilobytes (e.g. $32$ KB) and hence $21$ KB is considered as a significant saving. Also, the memory savings do not affect the run-time since  the controller synthesis and the corresponding mixed precision computations are all performed offline.}\\
In summary, our contributions are:
\begin{itemize}
\item We provide an end-to-end automatic tuning algorithm for robust explicit MPC that ensures satisfaction of
hard constraints on the state despite numerical errors;
\item We design a static and scalable error analysis algorithm for piecewise affine functions to bound
errors in controller implementations; and
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}

>>>>

In summary, our algorithm automatically and soundly synthesizes a suitable fixed-point mixed-precision implementation
for explicit robust MPC controllers.

@@

For smaller examples, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few hours.
In absolute terms, the memory saving corresponds to 21KB over a uniform
precision implementation on our largest benchmark. 
%\Mahmoud{It should be noticed that these savings are not affecting the run-time. Synthesis of the controller and the corresponding precisions are all performed offline.}
A large class of EMPC applications implement the controller in processors with limited computational power and memory;
typically,  these processors have memory in the order of tens of kilobytes (e.g., $32$ KB).
Thus, $21$ KB is a significant saving. 
Also, the memory savings do not affect the run-time performance 
since our controller synthesis and the corresponding mixed precision computations are all performed offline.

In summary, our contributions are:
\begin{itemize}
\item We provide an end-to-end automatic tuning algorithm for robust explicit MPC that ensures satisfaction of
hard constraints on the state despite numerical errors;
\item We design a static and scalable error analysis algorithm for piecewise affine functions to bound
errors in controller implementations; and
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}
